---
title: "Add weather variable"
author: "Qianqian Du"
date: "2023-07-12"
output: html_document
---


# Objective:

+ Get the weather data from daymet.
  + N
+ Incorporate the weather data into the corn data so that all variables can be included in GAM analysis.
  + get monthly precipitation amound for Apr to Sep
  + get the number of days that temperture exceed 32 for Apr to Sep
  + 
  
```{r echo=FALSE, eval=TRUE, message=FALSE, cache=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(ggplot2)
library(here)
library(patchwork)
library(data.table)
library(flextable)
library(plm)
library(gam)
library(mgcv)
library(aod)
library(glmmTMB)
library(matrixcalc)
library(gamm4)
library(gratia)
library(lme4)
library(insight)
library(scam)
library(forstringr)
```


```{r}
SD <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_CLC_2012_2022.csv") %>% 
  mutate(area = "SD")

IL_N <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_N_IL_2012_2022_1.csv") %>% 
  mutate(area = "IL_N")


IL_C <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_C_IL_2012_2022.csv") %>% 
  mutate(area = "IL_C")

IL_S <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_S_IL_2012_2022.csv") %>% 
  mutate(area = "IL_S")


OH <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_OH_2012_2022.csv") %>% 
  mutate(area = "OH")

Rendel_NofMineralView <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/Rendel_NofMineralView_2012_2022.csv") %>% 
  mutate(area = "Rendel_NofMineralView")

M01 <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_M01_2012_2022.csv") %>% 
  mutate(area = "M01")

DuraClub <- read.csv("/Users/qianqiandu/Library/CloudStorage/Box-Box/DIFM_HQ/Codes/Codes_indiv/QQ/Aggregate_reports/Data/field_weather_DuraClub_2012_2022.csv") %>% 
  mutate(area = "Duraclub")


```


```{r}
weather_data <- rbind(IL_N, IL_C) %>% 
  rbind(., IL_S) %>% 
  rbind(., OH) %>% 
  rbind(., Rendel_NofMineralView) %>% 
  rbind(., M01) %>% 
  rbind(., DuraClub)

```

```{r}

#/*=================================================*/
#' # ### === get precipitation data === ###
#/*=================================================*/

weather_data_analysis_precip <- weather_data %>% 
  mutate(month = case_when(yday <= 31 ~ "Jan",
                           yday > 31 & yday <= 59 ~ "Feb",
                           yday > 59 & yday <= 90 ~ "Mar",
                           yday > 90 & yday <= 120 ~ "Apr",
                           yday > 120 & yday <= 151 ~ "May",
                           yday > 151 & yday <= 181 ~ "Jun",
                           yday > 181 & yday <= 211 ~ "Jul",
                           yday > 211 & yday <= 240 ~ "Aug",
                           yday > 240 & yday <= 271 ~ "Sep",
                           yday > 271 & yday <= 301 ~ "Oct",
                           yday > 301 & yday <= 331 ~ "Nov",
                           yday > 331 & yday <= 366 ~ "Dec",
  )) %>% 
  data.table() %>% 
  .[year == 2021 | year == 2022, ] %>% 
  .[, .(month_precip = sum(prcp)), by = c("year", "month", "area")] %>% 
  dcast(., year + area ~ month, value.var = "month_precip")


weather_data_analysis_precip$year <- as.character(weather_data_analysis_precip$year)

#/*=================================================*/
#' # ### === get temperature data === ###
#/*=================================================*/

weather_data_analysis_temp <- weather_data %>% 
  mutate(month = case_when(yday <= 31 ~ "Jan",
                           yday > 31 & yday <= 59 ~ "Feb",
                           yday > 59 & yday <= 90 ~ "Mar",
                           yday > 90 & yday <= 120 ~ "Apr",
                           yday > 120 & yday <= 151 ~ "May",
                           yday > 151 & yday <= 181 ~ "Jun",
                           yday > 181 & yday <= 211 ~ "Jul",
                           yday > 211 & yday <= 240 ~ "Aug",
                           yday > 240 & yday <= 271 ~ "Sep",
                           yday > 271 & yday <= 301 ~ "Oct",
                           yday > 301 & yday <= 331 ~ "Nov",
                           yday > 331 & yday <= 366 ~ "Dec",
  )) %>% 
  data.table() %>%
  .[year == 2021 | year == 2022, ] %>%
  .[tmax >= 32, .(counts = .N), by = c("year", "month", "area")] %>% 
  dcast(., year + area ~ month, value.var = "counts") %>% 
  setnames("May", "May_t") %>%
  setnames("Jun", "Jun_t") %>% 
  setnames("Jul", "Jul_t") %>% 
  setnames("Aug", "Aug_t") %>% 
  setnames("Sep", "Sep_t")

weather_data_analysis_temp$year <- as.character(weather_data_analysis_temp$year) 

weather_data_analysis_temp[is.na(weather_data_analysis_temp)] <- 0


#/*=================================================*/
#' # ### === get Extreme Degree Days data === ###
#/*=================================================*/

weather_data_analysis_EDD <- weather_data %>% 
  mutate(month = case_when(yday <= 31 ~ "Jan",
                           yday > 31 & yday <= 59 ~ "Feb",
                           yday > 59 & yday <= 90 ~ "Mar",
                           yday > 90 & yday <= 120 ~ "Apr",
                           yday > 120 & yday <= 151 ~ "May",
                           yday > 151 & yday <= 181 ~ "Jun",
                           yday > 181 & yday <= 211 ~ "Jul",
                           yday > 211 & yday <= 240 ~ "Aug",
                           yday > 240 & yday <= 271 ~ "Sep",
                           yday > 271 & yday <= 301 ~ "Oct",
                           yday > 301 & yday <= 331 ~ "Nov",
                           yday > 331 & yday <= 366 ~ "Dec",
  )) %>% 
  data.table() %>%
  .[year == 2021 | year == 2022, ] %>% 
  .[, EDD_daily := pmax(0, tmax - 29)] %>% 
  .[month == "Apr"| month == "May"| month == "Jun" | month == "Jul"| month == "Aug"| month == "Sep", .(EDD = sum(EDD_daily)), by = c("year", "area")]

weather_data_analysis_EDD$year <- as.character(weather_data_analysis_EDD$year) 


weather_data_analysis_EDD$EDD %>% unique() %>% sort()

weather_data_analysis_EDD[EDD == "33.16", ]


```


```{r}
data_2021 <- readRDS("/Users/qianqiandu/Library/CloudStorage/Box-Box/Field_fixed_effects/Shared/Data/Data_for_analysis/data2021.rds")

data_2022 <- readRDS("/Users/qianqiandu/Library/CloudStorage/Box-Box/Field_fixed_effects/Shared/Data/Data_for_analysis/data2022.rds")

all_data <- rbind(data_2021, data_2022)

head(all_data)

# all_data$field %>% unique()

###--- create mean variables for the Mundlak method ---###

all_data <- all_data %>% 
  data.table() %>% 
  .[, mean_elev := mean(elev, na.rm = TRUE), by = field] %>% 
  .[, mean_slope := mean(slope, na.rm = TRUE), by = field] %>% 
  .[, mean_curv := mean(curv, na.rm = TRUE), by = field] %>% 
  .[, mean_tpi := mean(tpi, na.rm = TRUE), by = field] %>% 
  .[, mean_N := mean(n_rate, na.rm = TRUE), by = field]

all_data <- all_data %>% 
  mutate(area = case_when(field == "F98" ~ "OH",
                          field == "F104" ~ "OH",
                          field == "LaueLib80" ~ "IL_C",
                          field == "Dougshome" ~ "IL_N",
                          field == "NofMineralView" ~ "Rendel_NofMineralView", 
                          field == "Florence80" ~ "IL_N", 
                          field == "42ac" ~ "Rendel_NofMineralView", 
                          field == "DamgaardWest" ~ "IL_N",
                          field == "SyfordNorthEast" ~ "IL_C",
                          field == "JensenWest" ~ "IL_C",
                          field == "DJWest" ~ "IL_N", 
                          field == "Goldenrod" ~ "IL_S", 
                          field == "Field32" ~ "IL_C", 
                          field == "Elsie" ~ "IL_C", 
                          field == "Loveless" ~ "IL_C",
                          field == "Wirth" ~ "IL_N", 
                          field == "Rishel" ~ "IL_N", 
                          field == "F17" ~ "OH", 
                          field == "Andrews" ~ "IL_C",
                          field == "PerrySouth" ~ "IL_N", 
                          field == "Field1" ~ "IL_C",
                          field == "ShellebargerPainterCreekArcanum" ~ "OH", 
                          field == "M01" ~ "M01", 
                          field == "BostonWest" ~ "Rendel_NofMineralView",
                          field == "P1" ~ "Duraclub", 
                          field == "R37" ~ "Duraclub", 
                          field == "D09" ~ "Duraclub"
                          ))

```


```{r}

analysis_data <- left_join(all_data, weather_data_analysis_precip, by = c("area", "year")) %>% 
  left_join(., weather_data_analysis_temp, by = c("area", "year")) %>% 
  left_join(., weather_data_analysis_EDD, by = c("area","year"))


# saveRDS(analysis_data, "/Users/qianqiandu/Library/CloudStorage/Box-Box/Field_fixed_effects/Shared/Data/Data_for_analysis/field_with_weather.rds")



final_data <- analysis_data[, c("field", "year", "yield", "n_rate", "elev", "slope", "curv", "area", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "EDD")] %>% 
  na.omit()



```


# Run regression using GAM with precipitation and tempreature 


# Pooled 

```{r cache=TRUE}
pooled_gam <- gamm4(yield ~ s(n_rate, k = 4) + s(elev) + s(slope) + s(curv) + s(Apr) + s(May) + s(Jun) + s(Jul) + s(Aug) + s(Sep) + s(EDD), data = analysis_data)

summary(pooled_gam$gam)

```


# Fixed effect (field)

```{r cache=TRUE}
###--- gamm4() with field fixed effect ---###

fixed_gam <- gamm4(yield ~ s(n_rate, k = 4) + s(elev) + s(slope) + s(curv) + s(Apr) + s(May) + s(Jun) + s(Jul) + s(Aug) + s(Sep) + s(EDD) + field, data = final_data)

summary(fixed_gam$gam)

```


## Visualize the two models' production function 

```{r cache=TRUE}
##-- plot for without fixed effect --##
eval_data <- expand_grid(
  n_rate = seq(
    quantile(analysis_data$n_rate, prob = 0.10, na.rm = TRUE), 
    quantile(analysis_data$n_rate, prob = 0.90, na.rm = TRUE), 
    length = 300
  ),
  elev = mean(analysis_data$elev, na.rm = TRUE),
  slope = mean(analysis_data$slope, na.rm = TRUE),
  curv = mean(analysis_data$curv, na.rm = TRUE),
  Apr = mean(analysis_data$Apr, na.rm = TRUE),
  May = mean(analysis_data$May, na.rm = TRUE),
  Jun = mean(analysis_data$Jun, na.rm = TRUE),
  Jul = mean(analysis_data$Jul, na.rm = TRUE),
  Aug = mean(analysis_data$Aug, na.rm = TRUE),
  Sep = mean(analysis_data$Sep, na.rm = TRUE),
  EDD = mean(analysis_data$EDD, na.rm = TRUE)
) %>% 
data.table() 

##-- plot for without fixed effect --##
corn_price = 4
N_price = 0.4

pooled <- eval_data %>% 
  .[, y_hat := predict(pooled_gam$gam, newdata = eval_data)] %>% 
  .[, profit := y_hat*corn_price - n_rate*N_price]

EONR_pooled <- filter(pooled, profit == max(profit))


pooled_fig <- ggplot(data = pooled)+
  geom_line(aes(x = n_rate, y = y_hat))


##-- fixed effect --##

eval_data_fixed <- expand_grid(
  n_rate = seq(
    quantile(all_data$n_rate, prob = 0.05, na.rm = TRUE), 
    quantile(all_data$n_rate, prob = 0.95, na.rm = TRUE), 
    length = 300
  ),
  elev = mean(all_data$elev, na.rm = TRUE),
  slope = mean(all_data$slope, na.rm = TRUE),
  curv = mean(all_data$curv, na.rm = TRUE),
  field = "SyfordNorthEast"
) %>% 
data.table() 

##-- plot for with fixed effect --##

fixed <- eval_data_fixed %>% 
  .[, y_hat := predict(fixed_gam$gam, newdata = eval_data_fixed)]%>% 
  .[, profit := y_hat*corn_price - n_rate*N_price]

EONR_fixed <- filter(fixed, profit == max(profit))


fixed_fig <- ggplot(data = fixed)+
  geom_line(aes(x = n_rate, y = y_hat))


##-- over all figure --##
ggplot()+
  geom_line(aes(x = n_rate, y = y_hat), data = fixed, color = "red")+
  geom_line(aes(x = n_rate, y = y_hat), data = pooled)+
  theme_bw()+
  ggtitle("black is pooled")+
  ylab("Yield (bu/ac)")+
  xlab("N rate (lb/ac)")

```

Change the range of the figure
```{r cache=TRUE}

fixed_change_y <- eval_data_fixed %>% 
  .[, y_hat := predict(fixed_gam$gam, newdata = eval_data_fixed)] %>% 
  .[, y_hat_1 := y_hat - 330 ]


fixed_fig_1 <- ggplot()+
  geom_line(aes(x = n_rate, y = y_hat_1), data = fixed_change_y, color = "red")+
  geom_line(aes(x = n_rate, y = y_hat), data = pooled)+
  theme_bw()+
  ylab("Yield (bu/ac)")+
  xlab("N rate (lb/ac)")
  # annotate("text", label = "black is pooled, \n s(N, k = 6)", x = 130, y = 180)+
  # annotate("text", label = "pooled 231, fixed 250 (corner solu)", x = 200, y = 165)

fixed_fig_1

```


# Based on the random effect result, do the Mundlak test

## Random effect (with field means for the Mundlak method)

```{r cache=TRUE, cache.lazy = FALSE, eval=FALSE, echo=TRUE}
###--- gamm4() with random effect ---###

gamm4_random <- gamm4(yield ~ s(n_rate, k = 4) + s(elev, k = 8) + s(slope, k = 8) + s(curv, k = 8) + s(mean_N, k = 8) + s(mean_elev, k = 8) + s(mean_slope, k = 8) + s(mean_curv, k = 8), random = ~ (1 | field), data = all_data)
# takes 27 mins

# gamm4_random <- gamm4(yield ~ s(n_rate, k = 6) + s(elev, k = 6) + s(slope, k = 6) + s(curv, k = 6) + s(mean_N, k = 6) + s(mean_elev, k = 6) + s(mean_slope, k = 6) + s(mean_curv, k = 6), random = ~ (1 | field), data = all_data)

summary(gamm4_random$gam)
```


## Wald test for GAM model

$$
W = (R \hat{\beta} - r)' \ [\hat{\sigma}^2 R(X'X)^{-1}R']^{-1} \ (R \hat{\beta} - r) \sim \chi_{(r)}^2
$$

```{r cache=TRUE, eval=FALSE, echo=TRUE}
### Get the coefficient matrix of the gam model
coefficients <- coefficients(gamm4_random$gam) %>% as.matrix()
length(coefficients)

b <- coefficients[26:53, ]

b = b %>% as.matrix() #(20*1)
# b

### Generate the Identity matrix that is (rk * rk)
R = diag(length(b)) #(20*20)
# R 

# R_T <- t(R) since R is indentiy matrix, the transpose should be the same 

### Generate the Null hypothesis estimator 
r = matrix(0, length(b), 1) #(20*1)
# r

### Get the variance covariance matrix of the gam model
variance_covariance <- vcov(gamm4_random$gam)

variance_cov <- variance_covariance[26:53, 26:53] #(20 *20)
# variance

### Wald test step by step ###

## numerator 
Rb <- R %*% b

Rb_r <- Rb - r

RB_r_trans <- t(Rb_r)

numerator <- Rb_r %*% RB_r_trans

## denominator
R_var <- R %*% variance_cov

R_var_R <- R_var %*% R

inverse_var <- solve(R_var_R)

#system is computationally singular: reciprocal condition number = 1.03534e-17

## Wald statistic 

Wald <- RB_r_trans %*% inverse_var %*% Rb_r

Wald

critical_value <- qchisq(0.05, , lower.tail=FALSE)

critical_value

Wald>critical_value

```

